---
title: "p8105_hw1_as5457"
author: "Sunny Siddique"
date: "September 18, 2018"
output: html_document
---

# Problem 1


```{r random_sample}
library (tidyverse)

set.seed(1)

prob1_df = tibble(
  rand_samp = runif(10, min = 0, max = 5),
  vector_logical = rand_samp > 2,
  vector_char = c('Hi', 'my', 'name', 'is', 'Sunny', 'and', 'I', 'love', 'chicken', 'wings'),
  vector_factor = factor(c('male', 'male', 'male', 'male', 'male', 'female', 'female', 'female', 'female', 'female'))
)


```
```{r checking_classes}
class(prob1_df$vector_logical)
class(prob1_df$vector_char)
class(prob1_df$vector_factor)
```



```{r eval=FALSE}
data(prob1_df)
mean(prob1_df[["rand_samp"]])
mean(prob1_df[["vector_logical"]])
mean(prob1_df[["vector_char"]])
mean(prob1_df[["vector_factor"]])
```


When taking the means of the variables, I was able to retrieve means of the random sample and logical vector. The character and factor vectors did not retrieve means because they are neither numeric, nor logical.



```{r}
data(prob1_df)
as.numeric(prob1_df[["vector_logical"]])
as.numeric(prob1_df[["vector_char"]])
as.numeric(prob1_df[["vector_factor"]])
```
When the logical vector was coerced into being numeric, the numeric conversion led to values that were false (<2) to be assigned a value of 0 and values that were true (>2) to be assigned a value of 1. 

When the character vector was coerced into being numeric, the numeric conversion failed.

When the factor vector was coereced into being numeric, the conversion led to males being assigned a value of 2 and females being assigned a value of 1. 

In a second code chunk convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?

```{r}
as.factor(prob1_df[["vector_char"]])
as.numeric(prob1_df[["vector_char"]])

as.character(prob1_df[["vector_factor"]])
as.numeric(prob1_df[["vector_factor"]])
```

When I convered the character variable from character to factor to numeric, the conversion failed. However, when I convered the factor variable from factor to character to numeric, I was able to retrieve the same numeric assignments that I had received when I converted my factor variable directly to numeric. 



