---
title: "p8105_hw1_as5457"
author: "Sunny Siddique"
date: "September 18, 2018"
output: html_document
---

# Problem 1


```{r random_sample}
library (tidyverse)

set.seed(1)

prob1_df = tibble(
  rand_samp = runif(10, min = 0, max = 5),
  vector_logical = rand_samp > 2,
  vector_char = c('Hi', 'my', 'name', 'is', 'Sunny', 'and', 'I', 'love', 'chicken', 'wings'),
  vector_factor = factor(c('male', 'male', 'male', 'male', 'male', 'female', 'female', 'female', 'female', 'female'))
)

```
```{r checking_classes}
class(prob1_df$vector_logical)
class(prob1_df$vector_char)
class(prob1_df$vector_factor)
```



```{r eval=FALSE}

data(prob1_df)
mean(prob1_df[["rand_samp"]])
mean(prob1_df[["vector_logical"]])
mean(prob1_df[["vector_char"]])
mean(prob1_df[["vector_factor"]])

```

When taking the means of the variables, I was able to retrieve means of the random sample and logical vector. The character and factor vectors did not retrieve means because they are neither numeric, nor logical. The mean function in R only takes the mean of numeric and logical vectors. Therefore, we were unable to compute the mean for the character and factor variables. 



```{r eval = FALSE}
data(prob1_df)
as.numeric(prob1_df[["vector_logical"]])
as.numeric(prob1_df[["vector_char"]])
as.numeric(prob1_df[["vector_factor"]])
```
When the logical vector was coerced into being numeric, the numeric conversion led to values that were false (<2) to be assigned a value of 0 and values that were true (>2) to be assigned a value of 1. 

When the character vector was coerced into being numeric, the numeric conversion failed. This may be because there are no patterns in the character variables. 

When the factor vector was coereced into being numeric, the conversion led to males being assigned a value of 2 and females being assigned a value of 1. 

In a second code chunk convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?

```{r}
as.numeric(as.factor(prob1_df[["vector_char"]]))

as.numeric(as.character(prob1_df[["vector_factor"]]))
```

When we converted character to factor to numeric, each variable was assigned a new number. 
Second one doesn't work because we can't convert character to numeric. 


#Problem 2

This problem focuses on plotting and the use of inline R code.

Create a data frame comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector indicating whether the x + y > 0
A numeric vector created by coercing the above logical vector
A factor vector created by coercing the above logical vector

```{r}
set.seed (1)

prob2_df = tibble(
  x = rnorm (1000, mean = 0, sd = 1),
  y = rnorm (1000, mean = 0, sd = 1),
  logic_xy = (x + y > 0)
  #num_xy = as.numeric(prob2_df[["logic_xy"]]),
  #fac_xy = as.factor(prob2_df[["logic_xy"]])
)

prob2_df$num_xy = as.numeric(prob2_df[["logic_xy"]])
prob2_df$fac_xy = as.factor(prob2_df[["logic_xy"]])
```

##Write a short description of your vector using inline R code, including: the size of the dataset, the mean and median of x, the proportion of cases for which the logical vector is true.

The number of rows in this data frame is `r nrow(prob2_df)` and the number of columns in this data frame is `r ncol(prob2_df)`. The mean of the variable x is `r mean(prob2_df[["x"]])` and median of variable x is `r median(prob2_df$x)`. The proportion of cases for which the logical vector is true is `r mean(prob2_df$num_xy)`.


##Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

```{r}
library(ggplot2)
color_xy = ggplot(prob2_df, aes(x = x, y = y, color= logic_xy)) + geom_point()
color_xy

```

```{r}
color_numeric = ggplot(prob2_df, aes(x = x, y = y, color= num_xy)) + geom_point()
color_numeric
color_factor = ggplot(prob2_df, aes(x = x, y = y, color= fac_xy)) + geom_point()
color_factor
```


Exporting first scatterplot using ggsave.

```{r}
ggsave("scatter_plot.pdf", plot = color_xy)
```








